\bf A Hierarchical Bayesian Approach for Modeling Infant-Mortality and Wearout Failure Modes

### 2017-08-23 paper/paper.tex
Small edits.
Edits to abstract and intro

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf,float, bm}
\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
\usepackage{enumerate}
\usepackage[numbers]{natbib}
\usepackage{url} % not crucial - just used below for the URL

@@

\bigskip
\begin{abstract}
% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
When conducting lifetime analysis with field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure. The causes of these, while not interesting to the consumer per se, can lead to various observed lifetime distributions. Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate. 
Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision. These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model \citep{chan}. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to "infant mortality" and failure due to wearout. We fit the GLFP model using a hierarchical modeling approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using samples from the posterior distribution.
\end{abstract}

@@

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring early ("infant mortality") due to manufacturing defects.  After this "burn-in" period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability. This highlights the importance of choosing a sufficiently flexible model.

@@

Often no information is available to identify why a unit failed. This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. Where this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes "soft" constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

The structure of the paper is as follows.  Section~\ref{sec:Background} describes some of the potential applications of our method and summarizes previous related work found in the literature. Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates an application in an example. Section~\ref{sec:Hierarchical GLFP model} introduces a hierarchical model for extending the GLFP model to multiple populations. In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set. Subsection~\ref{sec:Data} describes and summarizes the Backblaze Hard Drive data set.  Subsection~\ref{sec:Model Comparisons} shows the use of approximate leave-one-out cross-validation as a way to do model selection between different hierarchical model specifications. We also address computational issues and modeling decisions that lead to a final model.  In subsection~\ref{sec:Drive-model Comparisons} we present an analysis comparing drive-model lifetime distributions.  Finally, in section~\ref{sec:Conclusions}, we discuss the strengths and limitations of our method and consider extensions and alternatives.

@@

% \bibliographystyle{plainnat}
\bibliographystyle{apalike}
\bibliography{sample}

>>>>

\documentclass[12pt]{article}
\usepackage{amsmath}
\usepackage{graphicx,psfrag,epsf,float, bm}
\graphicspath{C:/Users/Colin/Documents/GitHub/BB_data_analysis/paper/fig/}
\usepackage{enumerate}
\usepackage{natbib}
\usepackage{url} % not crucial - just used below for the URL

@@

\bigskip
\begin{abstract}
% Consumers have an interest in accurate assessment of product reliability.  However, lifetime data is often sparse and has limited information.  
When analyzing field data on consumer products, model-based approaches to inference require a model with sufficient flexibility to account for multiple kinds of failure. The causes of failure, while not interesting to the consumer per se, can lead to various observed lifetime distributions. Because of this, standard lifetime models, such as Weibull or lognormal may be inadequate. 
Furthermore, when the information carried by lifetime data are limited by sample size, censoring and truncation, estimates can be unstable and suffer from imprecision. These limitations are typical; for example, lifetime data for high-reliability products will naturally tend to be right-censored.

In this paper we present a method for joint estimation of multiple lifetime distributions based on the Generalized Limited Failure Population (GLFP) model. This 5-parameter model for lifetime data accommodates lifetime distributions with multiple failure modes:  early failure due to ``infant mortality'' and failure due to wearout. We fit the GLFP model using a hierarchical modeling approach.  Borrowing strength across populations, our method enables estimation with uncertainty of lifetime distributions even in cases where the number of model parameters is larger than the number of observed failures.  Moreover, using our Bayesian method, comparison of different product brands is straightforward because estimation of arbitrary functionals are easily computable using draws from the joint posterior distribution of the model parameters.
\end{abstract}

@@

The reliability of many engineered products follow a similar pattern. Relatively high rates of failure occurring early (``infant mortality'') due to manufacturing defects.  After this ``burn-in'' period, failure rates stabilize as the majority of defective units have failed and are no longer in the population.  Finally, after prolonged use, rates of failure increase due to wearout.  Ignoring this pattern in failure rates can lead to suboptimal decisions and spurious inferences about a product's reliability. This highlights the importance of choosing a sufficiently flexible model.

@@

Often no information is available to identify why a unit failed and even when the failed units are available, physical failure mode analysis (sometimes referred to as ``autopsy'') can be extremely time consuming and expensive. This presents a dilemma for analysts: the data may indicate multiple failure modes, contraindicating the use of a unimodal parametric distribution. However, without any knowledge of cause of failure, traditional competing risk models are not identifiable. \citet{chan} proposed the Generalized Limited Failure Population (GLFP) model. This model can provide a solution to the problem of unknown cause of failure (masking) in the special case where there is evidence of two modes of failure which impact different stages of product life. It avoids non-identifiability by introducing a parameter representing the population proportion defective. Where this parameter is zero, the GLFP model reduces to a unimodal distribution. \\

The GLFP model has a meaningful parameterization and accommodates lifetime data with infant mortality and wearout.  Unfortunately, it can require a lot of data to fit due to the model's complexity. When multiple sources of information are available, partial pooling, accomplished via hierarchical modeling, can reduce the amount of data required to produce stable parameter estimates. When multiple populations are of interest, comparisons based on separate, unrestricted GLFP model fits will be limited to those products with sufficient data. As we will show, hierarchical modeling of the GLFP parameters allows for borrowing of information across populations which imposes ``soft'' constraints on model parameters. This enables estimation of the lifetime distribution for all populations via shrinkage toward a "pooled" model, with the degree of shrinkage inversely related to the amount of population specific information. We demonstrate computation of various quantities of interest while comparing reliability across populations using numerical integration over the posterior distribution with samples obtained via  Markov Chain Monte Carlo.

The structure of the paper is as follows. In Section~\ref{Motivating example}, we describe, as a motivating example, the Backblaze Hard Drive data set. In Section~\ref{sec:Background}, we introduce our proposed method and potential applications and examine previous related work in the literature. Section~\ref{sec:GLFP model} introduces the GLFP model and illustrates its use with a subset of the Backblaze data. In section~\ref{sec:Hierarchical GLFP model} we discuss using hierarchical modeling of the GLFP model parameters to extend the GLFP model to multiple populations. In section~\ref{sec:Data analysis} we illustrate applying the hierarchical GLFP model to a real data set. First, we do model selection among four models of increasing complexity, using approximate leave-one-out cross-validation. Next, for the selected model, we assess the model fit using a graphical comparison of replicated data sets generated from the posterior distribution to the observed data. Finally, we present an evaluative comparison of populations taking into account practical considerations related to product lifetime. Finally, in section~\ref{sec:Conclusions}, we discuss the strengths and limitations of our method and consider extensions and alternatives.

@@

% \bibliographystyle{plainnat}
\bibliographystyle{plain}
\bibliography{sample}
